MA4413 Tutorials for Week 9
Q1. An IT competency test, used for staff recruitment, is devised so as to give a normal distribution of scores with a mean of 100. A random sample of 49 experienced IT users  who are given the test achieve a mean score of 121 with a standard deviation of 14. 

•	Perform a hypothesis test to assess whether this group of IT Users is unusual (i.e. have a different mean from the general population).
•	Compute a 95% confidence interval for the group.

Q2. A claim has been made that the mean body temperature of healthy adults is equal to 98.6 degrees. 
A sample of 106 people has produced a mean body temperature of 98.2 degrees and a standard deviation of 0.62. Test the claim using a 0.05 significance.

Q3. A manufacturer of computer monitors  has for many years used a process giving a mean life of 4700 hours and a standard deviation of 1460 hours. 
A new process is tried to see if it will increase the life significantly. A sample of 100 monitors gave a mean life of 5000 hours.  
Does the new process make a difference at the 5% level of significance?

Q4. In a study of store checkout scanners, 1234 items were checked and 20 of them were overcharges.
Use a 0.05 significance level to test the claim that with scanners, 1% of sales are overcharges.

Q5. The quality control manager at the Telektronic Company considers the production of telephone answering machines to be ’out of control’ when the overall rate of defects exceeds 4%. 
Testing of a random sample of 150 machines revealed that 9 are defective. The production manager claims that production is not out of control and no corrective action is necessary. Use a 0.05 significance level to test the production manager’s claim.




Q6. In a study of company salaries, salaries paid by 2 different IT companies were randomly selected.
•	For 40 Deltatech employees the mean is €23,870 and the standard deviation is €2960. 
•	For 35 Echelon employees , the mean is €22,025 and the standard deviation is €3065. 

At the 0.05 level of significance, test the claim that Deltatech employees earn the same as their  Echelon counterparts.

Q7 Does it pay to take preparatory courses for standardised tests such as the Comptia Exams? 
Using the sample data in the following table, compute the case-wise differences, the mean of the case-wise differences and the standard deviation of the case wise differences for the following data set.

Student	A	B	C	D	E	F	G	H	I	J
Score Before	700	840	830	860	840	690	830	1180	930	1070
Score After	720	840	820	900	870	700	800	1200	950	1080















Q8. A study was carried out in which researchers collected crime data. Of those convicted of
arson, 50 were drinkers and 43 abstained. Of those convicted of fraud, 63 were drinkers and 144
abstained. Use a 0.01 level of significance to test the claim that the proportion of drinkers among
convicted arsonists is greater than the proportion of drinkers convicted of fraud.





Q10. A bank is concerned about the amount of debt being accrued by customers using its credit
cards. The board of directors voted to institute an expensive monitoring system if the mean for all
the bank’s customers is greater than €2000. The bank randomly selected 50 credit-card holders and
determined the amounts they charged. For this sample group, the mean is €2177 and the standard ‘
deviation is € 1257. Using a 0.05 level of significance,  test the claim that the mean amount charged
is greater than €2000. Will the monitoring system be implemented? 

