Lights question from last week
Choose operators
binomial distribution using formula
binomial distribution using tables
expected value variance question

geometric distrubution
suppose you are tring to plug in a usb key into a pc. Suppose the chances of you inserting it correctly are 50%, as you repeatedly forget which side you used last time. What is the probability that you succeed with the third attempt. What is the probability that you need three or more attempts.

Poisson formula

Poisson formula using tables

%----------------------------------------------------------------------------------------------%

Numbers
number systems
binary addition and subtraction
logarithms and exponential
absolute value floor and ceiling function

%----------------------------------------------------------------------------------------------%

\documentclass{beamer}

\usepackage{default}

\begin{document}
The normal yield value quoted for a bond is the flat yield which is the fixed coupon divided by the current bond price as a percentage. The Gross Redemption Yield (GRY) takes into account not only the flat yield but also any capital gain received by holding the bond to maturity.

If for example a bond has a maturity value of £100 in 2 years time but is bought now for £95, there will be a capital gain of 5\% over two years (approxiamtely 2.47\% per annum) to add to the flat yield, if the bond is held to maturity. If the flat yield is 6\% per annum then the Gross Redemption Yield is 8.47\% per annum.


\end{document}

%----------------------------------------------------------------------------------------------%

RSS Grad Dip 3 - Question 3B2
 
 

SPECIMEN A

SPECIMEN A
6. Consider the AR(2) model
 
 
\[Y_t={1\over 3}Y_{t-1}+{1\over 12}Y_{t-2}+ \epsilon_t\]

for a process {Y_t}, where {\epsilon_t} is a white noise process.($-\infty \leq t \leq \infty$)

(i) Find the roots of the autoregressive characteristic equation and check that the stationarity condition is satisfied.(4)
 
(ii) Find the Yule-Walker equations that are satisfied by the autocorrelation function t.(4)
 
(iii) Obtain the value of $\rho_1$        (3)
 
(iv) Show that a general expression for the autocorrelation function is given by
\[\rho_t={35\over 44}{1\over 12}^t +{9\over 44}{-1\over 6}^t
\]

where $\tau \geq 0$     (9)


<hr>

2009/GD5/Q8
 
8. Monthly magazines are often directed at special interests of their readers, such as the countryside, wildlife, sport, motor cars. The marketing manager for one of these magazines is trying to encourage advertisers to place advertisements in her magazine. She therefore wishes to collect information on the number and types of people who read the magazine, regularly or occasionally, and their interests.
 
(i) Paying particular attention to the problems of constructing a sampling frame, undercoverage, non-response and likely requirements of advertisers, recommend a method
for selecting a sample from the readership. (10)
 
(ii) Comment on the types of information the manager should aim to collect, in addition to standard questions about name, address, age, occupation, etc. (5)
 
(iii) How could the frame be used, and how could the results of the exercise be kept up to date, to satisfy the likely needs of the advertisers? (5)

%----------------------------------------------------------------------------------------------%


Simple Spreadsheet commands that can used to convert decimal numbers to binary or hexadecimal, and vice versa.

Bin2Dec
Dec2bin
hex2dec
dec2hex

%----------------------------------------------------------------------------------------------%

<hr>
Chi Square Expected values
find the following values
row total
column total
overall total
%----------------%
Chi Square Test Statistic
   contribution for each cell
%----------------%
Chi Square Critical Value
   Degrees of freedom
%----------------%
R Chi Square Test
chisq.test()
%-----------------%
R Prop.test
%-----------------%
5
Pearson Correlation SS identities
For Loop
model goodness of fit metrics
transport NWC
transport LC
transport VAM shadow costs
10
%------------------%
MM3CB
15 
Normal Z score
Normal Using tables
Normal Complement Rule
Normal Interval Rule 
Normal Symmetry Rule  


robust regression
 anova
 interaction plots

Logistic regression
 SA heart data
 
this data set is available in the elemstatlearn package
 fit a logistic regression model where the chd us the binary outcome, and all other variables are used as predictor variables
 
use the AIC criterion 
http://aimotion.blogspot.ie/2011/11/machine-learning-with-python-logistic.html?m=1
 
http://www.datarobot.com/blog/classification-with-scikit-learn/


------------------------
SHINY

Minimimal examples CB1 CB2
 Exercises
 Reactivity
 Conditional panelling
 HTML examples
 Mathjax and HTML


SAS 5 videos
 > Shiny Web Apps taxonomy map
 > Control Charts u charts
 > Operations Research Videos
 > Dublin R population dynamics

>
 > Data Analytics masters
 >
 > R Coursera ESL
 > Julia videoes london user group
 > sas videos
 > python probability talk
 >
 > hadoop rhadoop
 > predictive analytics
 >
 > pandas wes mckinney


Data Analytics masters
 
R 
Julia
 sas
 python
 
hadoop
 predictive anslytics
 
pandas
 practical data science


Attribute charts
 
Variable charts
 
multivariate charts
 
np charts
>
 > Part D checklist
 >
 > use of colours in plot

 > scatterplot abline
 residual plots
 cooks distance

Part D checklist
 
use of colours in plot
 scatterplot
 title and subtitle in plot
 Grubbs Test For Outlier
 log transformation
Set theory
 Set operations
 complement
 union intersection
 set difference
 venn diagrams
 8 Disjoint Regions
 membership tables
 
proof by truth tables
 AND
 OR
 NOT
 Set difference
 symmetric difference


We will use a simplistic system for interpreting significance values (i.e. p-values).
 
If a p value is less than 0.02 we reject the nyll hypothesis.
 If the p value is greater than 0.05 we fail to reject the null hypothesis
 If the pvalue us between the two thresholds then we deem the procedure to be inconclusive. 

The null hypothesis is that the true correlation coefficient is zero (which is to say, no linear relationship exists). 


Normal Distribution 
normal probability plot 

( qqplot boxplots histograms density plots)
 %-----------------------------------------------%
 skew and kurtosis describe function in the psych package 
grubbs test for outliers (outliers package)
 shapiro wilk test
 
anderson darling test (package normtest)
 
kolmogorov smirnov test for a single sample
 
correlation. Cor.test
 log transformation
 %----------------------------------------------%
 The t-tests 
variance test boxplots
 two sample tests 
paired t tests
 %-----------------------------------------------------%
 non parametric distributions
 ks-test
 wilcixon test for medians
Install.packages("packagename")
 library(packagename)
 #########
 data(datasetname) 
help(datasetname)
 
summary(datasetname)
 names(datasetname)
 
##########
 summary(datasetname$variablename)
 
## or 
attach(datasetname)
 summary(variablename)
 detach(datasetname)
 
##########
 
class(objectname)
 names(objectname)
 str(objectname)
 mode(objectname)
 
##########
 plot(variablename)
 plot(variablename1,variablename2)
 hist(variablename)


<hr>
Coursera data science specialization
 quality mgmt
 stata logistic regression
 susquehanna qr intern
 math biostats
 Dublin R







to Script 











ANOVA with R
 
The null hypothesis is that population mean is the same for each of the groups.
 
The alternative hypothesis is that there is at least  one group with a different population mean diffetent to the rest
 
the first tjing we do is to re-write the data in long form
 
Important :remember to construct tbe grouping variable as a factor. 


<hr>
Categorical variables are referred to by R (and in fact several branches of statistics as "factors". 

The categories for a factor are known as levels. 

Sex is an example of a factor variables, where the levels are Male and Female. 

Factors differ from character vectors. R woyld nit recognize a character variable as indicating membership of a category, rather it would just identify it as a piece of text..


 
-------------------------------------------

 
When computing confidence intervals or performing hypothesis tests, the most computationally complex component is the calculation of the standard error.
 
For each procedure, there is a standard error formula. These are usually given in an appendix to examination papers. It is strongly recommended that you familiarise yourself with these before your exam.



Scatterplots are useful for determining outliers. There is no single definition of an outlier, rather it depends on the type of analysis being used.


Medical statistic hypothesis testing
 
Type I error This describes a hypothesis test in which the null hypothesis was rejected, when it should not have been. 

Type II error This describes a hypothesis test in which the null hypothesis was rejected when it was in fact true.
 
-----------------------------------
 \begin{description}
 \item Specificity
 
\item Sensitivity
 \end{description}


Scatterplots are used to visualize the relationship between two continuous variables. The position of each point on the plot indicates the meaurement by both variables. Scatterplots are useful for determining the possible relationship between the variables.

Be mindful that the corration may be spurious. That is to say : The scatterplot indicates a relationship that does not really exist. Rather the points align in a linear manner on the scatterplot by chance. Special consideration should go into designing a scientific experiment to avoid such problems, including that there is a suitably large sample size. As the sample size gets less likely spurious correlation is less likely. 
