
\documentclass[a4]{beamer}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{newlfont}
\usepackage{amsmath,amsthm,amsfonts}
%\usepackage{beamerthemesplit}
\usepackage{pgf,pgfarrows,pgfnodes,pgfautomata,pgfheaps,pgfshade}
\usepackage{mathptmx}  % Font Family
\usepackage{helvet}   % Font Family
\usepackage{color}

\mode<presentation> {
 \usetheme{Default} % was
 \useinnertheme{rounded}
 \useoutertheme{infolines}
 \usefonttheme{serif}
 %\usecolortheme{wolverine}
% \usecolortheme{rose}
\usefonttheme{structurebold}
}

\setbeamercovered{dynamic}

\title[MA4102]{Business Maths 2 \\ {\normalsize MA4103 Tutorial}}
\author[Kevin O'Brien]{Kevin O'Brien \\ {\scriptsize Kevin.obrien@ul.ie}}
\date{Autumn Semester 2011}
\institute[Maths \& Stats]{Dept. of Mathematics \& Statistics, \\ University \textit{of} Limerick}

\renewcommand{\arraystretch}{1.5}
%---------------------------------------------------------------------%
\begin{frame}
\frametitle{Bivariate Data}
\begin{itemize}
\item Very often we are interested in the relationship between two
variables.
\item For example we might be interested in the relationship between
unemployment and inflation.
\item The investigation of a relationship between two variables
begins with an attempt to discover the approximate form of the
relationship by graphing the data using a scatter plot.
\end{itemize}	
\end{frame}
%---------------------------------------------------------------------%
\begin{frame}
\frametitle{Bivariate Data}
\begin{itemize}
\item  A scatter plot is a two-dimensional plot showing the (x,y) value for each
observation. 

\item  Using these plots, we can quickly determine whether there is
any pronounced relationship and if so, whether the relationship may be
treated as approximately linear.\\
Response variable Y (dependent variable).\\
Explanatory variable X (independent variable).
\item A response variable is the variable whose variation we wish to explain.
\item The explanatory variable is a variable used to explain the variation in the
response variable.
\end{itemize}	
\end{frame}
%---------------------------------------------------------------------%
\begin{frame}
\frametitle{Bivariate Data}
\begin{itemize}
\item The first step in determining the relationship between two
variables is to draw a scatter plot.
\item After establishing that a linear relationship exists between the
two variables X and Y, we need to measure the strength of this
relationship.
\item In order to do this, we need some measure of the association
or relationship between the two variables.
\item There are several ways of doing this - the most common
measure is the Pearson product moment correlation
coefficient usually known as the correlation coefficient.

\end{itemize}	
\end{frame}

%---------------------------------------------------------------------%
\begin{frame}
\frametitle{Properties of the Correlation Coefficient}
\begin{enumerate}
\item $-1 \leq r \leq +1$
\item r = +1 or -1 represents a perfect linear correlation or a perfect linear
relationship between the variables.
\item r = 0 indicates little or no relationship i.e. as X increases there is no
definite tendency for the value of Y to increase or decrease in a straight line.
\item r close to +1 indicates a large positive correlation i.e. Y tends to increase
as X increases.
\item r close to -1 indicates a large negative correlation i.e. Y tends to decrease
as X increases.
\item The more r differs from 0, the stronger the linear relationship between the
two variables.
\item The sign of r indicates the direction of the relationship.
\end{enumerate}
\end{frame}

%---------------------------------------------------------------------%
\begin{frame}
\frametitle{Properties of the Correlation Coefficient}
Example: We are given data for 6 graduates. Below is their
final QCA and their corresponding starting salary after
graduation.

% Subject 1 2 3 4 5 6
% Final QCA 2.8 3.4 3.2 3.8 3.2 3.4
% Starting Salary 20 000 24 500 23 000 25 000 20 000 22 500

Calculate the sample correlation coefficient.
\end{frame}


%---------------------------------------------------------------------%
\begin{frame}
\frametitle{Regression}
A statistical measure that attempts to determine the strength of the relationship between one dependent variable 
(usually denoted by Y) and a series of other changing variables (known as independent variables).

Regression takes a group of random variables, thought to be predicting Y, and tries to find a mathematical relationship between them. This relationship is typically in the form of a straight line (linear regression) that best approximates all the individual data points.

\end{frame}


%---------------------------------------------------------------------%
\begin{frame}
\frametitle{Regression}

Two events can consistently correlate with each other but not have any causal relationship. 
An example is the relationship between reading ability and shoe size across the whole population of the United States. 
If someone performed such a survey, they would find that the larger shoe sizes correlate with better reading ability, but 
this does not mean large feet cause good reading skills. Instead it's caused by the fact that young children have small 
feet and have not yet (or only recently) been taught to read. In this case, the two variables are more accurately correlated with a third: age.
\end{frame}
%---------------------------------------------------------------------%
\begin{frame}
\frametitle{Correlation Coefficient}

A correlation coefficient is a number between -1 and 1 which measures the degree to which two variables are linearly related. 

If there is perfect linear relationship with positive slope between the two variables, we have a correlation coefficient of 1; if there is positive correlation, whenever one variable has a high (low) value, so does the other. If there is a perfect linear relationship with negative slope between the two variables, we have a correlation coefficient of -1; if there is negative correlation, whenever one variable has a high (low) value, the other has a low (high) value. A correlation coefficient of 0 means that there is no linear relationship between the variables.

There are a number of different correlation coefficients that might be appropriate depending on the kinds of variables being studied.

\end{frame}
%---------------------------------------------------------------------%


\begin{frame}
\frametitle{Pearson's Product Moment Correlation Coefficient}

Pearson's product moment correlation coefficient, usually denoted by r, is one example of a correlation coefficient. It is a measure of the linear association between two variables that have been measured on interval or ratio scales, such as the relationship between height in inches and weight in pounds. 

However, it can be misleadingly small when there is a relationship between the variables but it is a non-linear one.

There are procedures, based on r, for making inferences about the population correlation coefficient. 
However, these make the implicit assumption that the two variables are jointly normally distributed. 
When this assumption is not justified, a non-parametric measure such as the Spearman Rank Correlation Coefficient might be more appropriate.

\end{frame}
%---------------------------------------------------------------------%


\begin{frame}
\frametitle{Least Squares}
The method of least squares is a criterion for fitting a specified model to observed data. 
For example, it is the most commonly used method of defining a straight line through a set of points on a scatterplot.
\end{frame}
%---------------------------------------------------------------------%
\begin{frame}
\frametitle{Regression Equation}
A regression equation allows us to express the relationship between two (or more) variables algebraically. It indicates the nature of the relationship between two (or more) variables. In particular, it indicates the extent to which you can predict some variables by knowing others, or the extent to which some are associated with others.



A linear regression equation is usually written
\[Y = a + bX + e\]
where
\begin{itemize}
\item Y is the dependent variable 
\item a is the intercept 
\item b is the slope or regression coefficient 
\item X is the independent variable (or covariate) 
\item e is the error term 
\end{itemize{
The equation will specify the average magnitude of the expected change in Y given a change in X.

The regression equation is often represented on a scatterplot by a regression line.
\end{frame}

%---------------------------------------------------------------------%
\begin{frame}
\frametitle{Regression Line}

A regression line is a line drawn through the points on a scatterplot to summarise the relationship between the variables being studied. When it slopes down (from top left to bottom right), this indicates a negative or inverse relationship between the variables; when it slopes up (from bottom right to top left), a positive or direct relationship is indicated.

The regression line often represents the regression equation on a scatterplot.

\end{frame}

%---------------------------------------------------------------------%
\begin{frame}
\frametitle{Types of Regression}

\begin{itemize}
\item \textbf{Simple Linear Regression}
Simple linear regression aims to find a linear relationship between a response variable and a possible predictor variable by the method of least squares.



\item \textbf{Multiple Regression}
Multiple linear regression aims is to find a linear relationship between a response variable and several possible predictor variables.



\item \textbf{Nonlinear Regression}
Nonlinear regression aims to describe the relationship between a response variable and one or more explanatory variables in a non-linear fashion.
\end{itemize}
\end{frame}
%---------------------------------------------------------------------%
\begin{frame}
\frametitle{Residual}
Residual (or error) represents unexplained (or residual) variation after fitting a regression model. 
It is the difference (or left over) between the observed value of the variable and the value suggested by the regression model.
\end{frame}

%---------------------------------------------------------------------%



%---------------------------------------------------------------------%



\end{document}