

\documentclass{beamer}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphics}

\begin{document}
	%-------------------------------------------------- %
	\begin{frame}
		\bigskip
		{
			\Huge
			\[ \mbox{Financial Mathematics}  \]
			\huge
			\[ \mbox{Fisher's Equation}  \]
		}
		
		{
			\LARGE
			\[ \mbox{www.MathsResource.com}  \]
			%\[ \mbox{Twitter: @StatsLabDublin} \]
		}
	\end{frame}


%===================================================================================================================%

%--- FISHERS EQUATION
%-------- http://en.wikipedia.org/wiki/Fisher_equation
 
The Fisher equation in financial mathematics and economics estimates the relationship between nominal and real interest rates inflation. 

It is named after Irving Fisher, who was famous for his works on the theory of interest. 

In finance, the Fisher equation is primarily used in YTM calculations of bonds or IRR calculations of investments. In economics, this equation is used to predict nominal and real interest rate behavior.
 
Letting r denote the real interest rate, i denote the nominal interest rate, and let $\pi$ denote the inflation rate, the Fisher equation is:
 
This is a linear approximation, but as here, it is often written as an equality:


 
The Fisher equation can be used in either ex-ante (before) or ex-post (after) analysis. Ex-post, it can be used to describe the real purchasing power of a loan:
 
Rearranged into an expectations augmented Fisher equation and given a desired real rate of return and an expected rate of inflation $\pi^e$ (with superscript e meaning "expected") over the period of a loan, it can be used as an ex-ante version to decide upon the nominal rate that should be charged for the loan:
 
This equation existed before Fisher, but Fisher proposed a better approximation which is given below. The approximation can be derived from the exact equation:
 


%===================================================================================================================%



%--------------------------------------------------
\begin{frame}
\frametitle{Fisher's Equation}
\Large
\begin{itemize}
\item The market rate of return on the 4.25\% UK government bond maturing on 8 March 2050 is 3.81\% per year. \item Let's assume that this can be broken down into a real rate of exactly 2\% and an inflation premium of 1.775\% \\ \textit{(no premium for risk, as government bond is considered to be "risk-free")}:
\end{itemize}
\[ 1.02 \times 1.01775 = (1 + 0.02) \times (1 + 0.01775) = 1.0381 \]

\end{frame}
%--------------------------------------------------
	\begin{frame}
		\frametitle{Fisher's Equation}
This article implies that you can ignore the least significant term in the expansion (0.02 × 0.01775 = 0.00035 or 0.035\%) and just call the nominal rate of return 3.775\%, on the grounds that that is almost the same as 3.81\%.
\end{frame}

%--------------------------------------------------
\begin{frame}
\frametitle{Fisher's Equation}
\Large
\begin{itemize}
\item At a nominal rate of return of 3.81\% pa, the value of the bond is £107.84 per £100 nominal. 
\item At a rate of return of 3.775\% pa, the value is £108.50 per £100 nominal, or 66p more.
\end{itemize}
\end{frame}

%--------------------------------------------------
\begin{frame}
	\frametitle{Fisher's Equation}
	\Large
\begin{itemize}
	\item
The average size of actual transactions in this bond in the market in the final quarter of 2005 was £10 million. 
\item So a difference in price of 66p per £100 translates into a difference of £66,000 per deal.
\end{itemize}
\end{frame}

\end{document}

Maths Resource

http://www.csusm.edu/mathlab/documents/M132BusCalcFormulas%20r1-12e.pdf
http://www.math.ubc.ca/~chau/elasticity.pdf
http://www.textbooksonline.tn.nic.in/books/12/std12-bm-em-1.pdf
