\documentclass{beamer}

\usepackage{amsmath}
\usepackage{amssymb}

\begin{document}


\begin{frame}
\Large
% FIX TITLE PAGE
\[\mbox{Annuities}\]
\end{frame}
%------------------------------------------------%
\begin{frame}
\frametitle{Annuities}
\begin{itemize}
\item A financial product sold by financial institutions that is designed to accept and grow funds from an individual and then, upon annuitization, pay out a stream of payments to the individual at a later point in time. Annuities are primarily used as a means of securing a steady cash flow for an individual during their retirement years.
 
\item Annuities can be structured according to a wide array of details and factors, such as the duration of time that payments from the annuity can be guaranteed to continue. Annuities can be created so that, upon annuitization, payments will continue so long as either the annuitant or their spouse is alive. Alternatively, annuities can be structured to pay out funds for a fixed amount of time, such as 20 years, regardless of how long the annuitant lives.
 
\item Annuities can be structured to provide fixed periodic payments to the annuitant or variable payments. The intent of variable annuities is to allow the annuitant to receive greater payments if investments of the annuity fund do well and smaller payments if its investments do poorly. This provides for a less stable cash flow than a fixed annuity, but allows the annuitant to reap the benefits of strong returns from their fund's investments.
 
 \end{itemize}
\end{frame}
%------------------------------------------------%
\begin{frame}
\frametitle{Annuities}
Formula for Finding the Periodic payment(R), Given A:
\[R = A/(1+(1-(1+(j/m) )^(-(n-1))/(j/m)\]
\end{frame}

%------------------------------------------------%
\begin{frame}
\frametitle{Annuities}
Examples:
Find the periodic payment of an annuity due of $\$70000$, payable annually for 3 years at 15\% compounded annually.
\[R = 70000/(1+[(1-(1+((.15)/1) )]^{(-(3-1))/((.15)/1))}\]
\[R = 70000/2.625708885\]
\[R = 26659.46724\]
\end{frame}

%------------------------------------------------%
\begin{frame}
\frametitle{Annuities}
Find the periodic payment of an annuity due of $250700, payable quarterly for 8 years at 5% compounded quarterly.
R= 250700/(1+〖(1-(1+((.05)/4) )〗^(-(32-1))/((.05)/4))
R = 250700/26.5692901
R = $9435.71
\end{frame}

%------------------------------------------------%
\begin{frame}
\frametitle{Annuities}
END PAGE SLIDE
\end{frame}
\end{document}

Maths Resource

http://www.csusm.edu/mathlab/documents/M132BusCalcFormulas%20r1-12e.pdf
http://www.math.ubc.ca/~chau/elasticity.pdf
http://www.textbooksonline.tn.nic.in/books/12/std12-bm-em-1.pdf
